\documentclass{propositions}

%% Turn off page numbering for the propositions and make the margins on both
%% sides equal and symmetrical.
\geometry{twoside=false}
\pagestyle{empty}

% \usepackage{fontspec}
% \defaultfontfeatures{Path = /usr/local/texlive/2024/texmf-dist/fonts/opentype/public/fontawesome/}
% \usepackage{fontawesome}

% \setmainfont[Path = fonts/libertinus/, ItalicFont=libertinusserif-italic.otf, BoldFont=libertinusserif-bold.otf, BoldItalicFont=libertinusserif-bolditalic.otf]{libertinusserif-regular.otf}
% \setsansfont[Path = fonts/libertinus/, BoldFont=libertinussans-bold.otf, ItalicFont=libertinussans-italic.otf]{libertinussans-regular.otf}
% \setmathfont[Path = fonts/libertinus/]{libertinusmath-regular.otf}
% \setmonofont[Scale=MatchLowercase, Path=fonts/inconsolata/]{Inconsolata-Regular.ttf}
%% The default style for text is Tahoma (sans-serif).
%\renewcommand*\familydefault{\rmfamily}

\begin{document}

%% Specify the title and author of the thesis. This information will be used on
%% both the English and Dutch side and in the metadata of the final PDF.
\title{Learning Event-Based Robot Vision}
\author{Jesse J.}{Hagenaars}

\begin{center}

{\Large\titlefont\bfseries Propositions}

\medskip

accompanying the dissertation

\medskip

%% Print the title.
{\makeatletter
\titlestyle\bfseries\large\@title
\makeatother}

%% Print the optional subtitle.
{\makeatletter
\ifx\@subtitle\undefined\else
    \titlefont\titleshape\@subtitle
\fi
\makeatother}

\smallskip

by

\smallskip

%% Print the full name of the author.
\makeatletter
{\large\titlefont\bfseries\@firstname\ {\titleshape\@lastname}}
\makeatother

\end{center}

\vspace{2.5pt}

\begin{enumerate}
\itemsep0em 
    \item To fully exploit the latency and efficiency benefits of neuromorphic hardware, all temporal integration of sensory information should happen inside the spiking neural network, not in its input encoding. (This dissertation)
    \item Self-supervised learning is essential for robust autonomous robots, enabling both large-scale pretraining and continual adaptation to distribution shifts when deployed in the real world. (This dissertation)
    \item Dedicated inertial sensing is not necessary for successful drone flight: an event camera combined with a recurrent neural network can replace the IMU in a drone's control loop. (This dissertation)
    \item To remain relevant, neuromorphic computing should focus on edge and robotic deployment rather than competing with the scaling trend toward ever-larger models and server infrastructure.
    \item Sensors that extend perception beyond conventional alternatives benefit disproportionately from self-supervised learning, because ground truth cannot be obtained where no reference sensor can reach.
    % \item Event cameras benefit disproportionately from self-supervised learning, because the high temporal resolution at which they operate makes supervised labels impractical to obtain at matching rates. (This dissertation)
    \item Constraining researchers in their available (GPU) compute is a powerful catalyst for deep learning engineering breakthroughs.
    \item Making a robot work reliably ten times in a row teaches you more than making ten different robots work once.
    % \item The most important skill a PhD develops is not domain expertise but the ability to learn and navigate unfamiliar fields independently.
    \item Congratulations! That you can read these words means you have won the \emph{Lottery of Life} at least thrice: fortune of birthplace, birth era, and biological endowment. Maintaining awareness of these unearned advantages is a foundation for contentment and a necessary corrective for the illusion that our outcomes are entirely self-made.
    \item The \emph{Golden Rule} ``Do unto others as you would have them do unto you'' is a potent local learning rule. However, as with local rules in neural networks, global feedback is required for globally optimal outcomes: the Golden Rule's adoption across many cultures has not prevented systemic exploitation between them.
    \item Current AI limitations preserve human cognition by requiring deep problem understanding for quality outputs. This inadequacy is our last barrier; future capable AI will trigger general cognitive decline through our innate efficiency-seeking.
\end{enumerate}

\vspace{2.5pt}

%% Apart from the name and title of the supervisor, the following text is
%% dictated by the promotieregelement.
\begin{center}
\noindent These propositions are regarded as opposable and defendable, and have been approved as such by the
promotors prof.\ dr.\ G.\ C.\ H.\ E.\ de Croon and \\prof.\ dr.\ S.\ M.\ Boht\'e.
\end{center}

%% \clearpage
%% {\selectlanguage{dutch}

%% \begin{center}

%% {\Large\titlefont\bfseries Stellingen}

%% \bigskip

%% behorende bij het proefschrift

%% \bigskip

%% %% Print the title.
%% {\makeatletter
%% \titlestyle\bfseries\large\@title
%% \makeatother}

%% %% Print the optional subtitle.
%% {\makeatletter
%% \ifx\@subtitle\undefined\else
%%     \titlefont\titleshape\@subtitle
%% \fi
%% \makeatother}

%% \bigskip

%% door

%% \bigskip

%% %% Print the full name of the author.
%% \makeatletter
%% {\large\titlefont\bfseries\@firstname\ {\titleshape\@lastname}}
%% \makeatother

%% \end{center}

%% \bigskip
%% \bigskip

%% \begin{enumerate}

%% \item Stelling 1.
%% \item Stelling 2.
%% \item Stelling 3.
%% \item Stelling 4.
%% \item Stelling 5.
%% \item Stelling 6.
%% \item Stelling 7.
%% \item Stelling 8.
%% \item Stelling 9.
%% \item Stelling 10.

%% \end{enumerate}

%% \bigskip
%% \bigskip

%% %% Apart from the name and title of the supervisor, the following text is
%% %% dictated by the promotieregelement.
%% \begin{center}
%% Deze stellingen worden opponeerbaar en verdedigbaar geacht en zijn als zodanig goedgekeurd door de promotoren prof.\ dr.\ A.\ van Deursen and dr.\ A.\ Zaidman.
%% \end{center}

%% }

\end{document}

