\chapter*{List of Publications}
\addcontentsline{toc}{chapter}{List of Publications}
\setheader{List of Publications}
\label{publications}

\noindent
$\dagger$~~Equal contribution.\\
\faFileTextO~~Included in this thesis.
\section*{Journal Papers}
\vspace{5pt}
\begin{etaremune}{\small
\item[\faFileTextO~~\hspace{4pt}7.] \emph{\textbf{F.\ Paredes-Vall\'es$^\dagger$}, J.\ J.\ Hagenaars$^\dagger$, J.\ D.\ Dupeyroux$^\dagger$, S.\ Stroobants, Y.\ Xu, G.\ C.\ H.\ E.\ de Croon}: Fully neuromorphic vision and control for autonomous drone flight. Under review at Science Robotics. 2023.
\item[\faFileTextO~~\hspace{4pt}6.] \emph{C.\ De Wagter$^\dagger$, \textbf{F.\ Paredes-Vall\'es}$^\dagger$, N.\ Sheth$^\dagger$, G.\ C.\ H.\ E.\ de Croon}: The sensing state-estimation and control behind the winning entry to the 2019 Artificial Intelligence Robotic Racing competition. Field Robotics. 2022.
\item[\faFileTextO~~\hspace{4pt}5.] \emph{C.\ De Wagter$^\dagger$, \textbf{F.\ Paredes-Vall\'es}$^\dagger$, N.\ Sheth$^\dagger$, G.\ C.\ H.\ E.\ de Croon}: Learning fast in autonomous drone racing. Nature Machine Intelligence (NMI). 2021.
\item[\faFileTextO~~\hspace{4pt}4.] \emph{D.\ B.\ de Jong, \textbf{F.\ Paredes-Vall\'es}, G.\ C.\ H.\ E.\ de Croon}: How do neural networks estimate optical flow? A neuropsychology-inspired study. IEEE Transactions on Pattern Analysis and Machine Intelligence (TPAMI). 2021.
\item[3.] \emph{J.\ J.\ Hagenaars, \textbf{F.\ Paredes-Vall\'es}, G.\ C.\ H.\ E.\ de Croon}: Evolved neuromorphic control for high speed divergence-based landings of MAVs. IEEE Robotics and Automation Letters (RA-L). 2020.
\item[\faFileTextO~~\hspace{4pt}2.] \emph{\textbf{F.\ Paredes-Vall\'es}, K.\ Y.\ W.\ Scheper, G.\ C.\ H.\ E.\ de Croon}: Unsupervised learning of a hierarchical spiking neural network for optical flow estimation: From events to global motion perception. IEEE Transactions on Pattern Analysis and Machine Intelligence (TPAMI). 2019.
\item[1.] \emph{S.\ Garc\'ia-Nieto, J.\ Velasco-Carrau, \textbf{F.\ Paredes-Vall\'es}, J.\ V.\ Salcedo, R. Simarro}: Motion equations and attitude control in the vertical flight of a VTOL bi-rotor UAV. Electronics. 2019.
}\end{etaremune}

\section*{Conference Papers}
\vspace{5pt}

\begin{etaremune}{\small
		\item[8.] \emph{Y.\ Wu, \textbf{F.\ Paredes-Vall\'es}, G.\ C.\ H.\ E.\ de Croon}: Lightweight event-based optical flow estimation via iterative deblurring. Under review at IEEE International Conference on Robotics and Automation (ICRA). 2024.
		\item[\faFileTextO~~\hspace{4pt}7.] \emph{\textbf{F.\ Paredes-Vall\'es}, K.\ Y.\ W.\ Scheper, C.\ De Wagter, G.\ C.\ H.\ E.\ de Croon}: Taming contrast maximization for learning sequential, low-latency, event-based optical flow. IEEE International Conference on Computer Vision (ICCV). 2023.
		\item[\faFileTextO~~\hspace{4pt}6.] \emph{R.\ J.\ Bouwmeester, \textbf{F.\ Paredes-Vall\'es}, G.\ C.\ H.\ E.\ de Croon}: NanoFlowNet: Real-time dense optical flow on a nano quadcopter. IEEE International Conference on Robotics and Automation (ICRA). 2023.
		\item[\faFileTextO~~\hspace{4pt}5.] \emph{J.\ J.\ Hagenaars$^\dagger$, \textbf{F.\ Paredes-Vall\'es}$^\dagger$, G.\ C.\ H.\ E.\ de Croon}: Self-supervised learning of event-based optical flow with spiking neural networks. Advances in Neural Information Processing Systems (NeurIPS). 2021.
		\item[4.] \emph{J.\ D.\ Dupeyroux, J.\ J.\ Hagenaars, \textbf{F.\ Paredes-Vall\'es}, G.\ C.\ H.\ E.\ de Croon}: Neuromorphic control for optic-flow-based landing of MAVs using the Loihi processor. IEEE International Conference on Robotics and Automation (ICRA). 2021.
		\item[\faFileTextO~~\hspace{4pt}3.] \emph{\textbf{F.\ Paredes-Vall\'es}, G.\ C.\ H.\ E.\ de Croon}: Back to event basics: Self-supervised learning of image reconstruction for event cameras via photometric constancy. IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR). 2021.
		\item[2.] \emph{J.\ J.\ Hagenaars, \textbf{F.\ Paredes-Vall\'es}, G.\ C.\ H.\ E.\ de Croon}: Evolved neuromorphic control for high speed divergence-based landings of MAVs. IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS). 2020.
		\item[1.] \emph{\textbf{F.\ Paredes-Vall\'es}, D.\ P.\ Magree, E.\ N.\ Johnson}: Direct feature correspondence in vision-aided inertial navigation for unmanned aerial vehicles. International Conference on Unmanned Aircraft Systems (ICUAS). 2017.
}\end{etaremune}