\chapter*{Summary}
\addcontentsline{toc}{chapter}{Summary}
\setheader{Summary}

\dropcap{V}{ision} is central to autonomous flight. From crop monitoring and infrastructure inspection to search-and-rescue in disaster zones and defense, the ability to perceive and navigate a complex environment in real time determines what a drone can do and where it can go. Current computer vision achieves impressive results, but at considerable cost: deep neural networks running on graphics processors consume tens of watts and add hundreds of grams. For larger drones, this limits flight time and the intelligence that can run on board. For the smallest platforms---those light enough to safely operate near people, navigate collapsed buildings, or monitor ecosystems in massive swarms---it is prohibitive altogether. Biology suggests that this trade-off is not inevitable: insects navigate complex environments with brains consuming microwatts of power, relying on sparse, event-driven neural computation rather than dense, synchronous processing.

This dissertation investigates how event-based vision can enable effective autonomous behavior on flying robots. Event cameras, inspired by the biological retina, respond asynchronously to brightness changes rather than capturing full frames, offering microsecond temporal resolution at milliwatts of power and high dynamic range. Neuromorphic processors extend this paradigm to computation through spiking neural networks that communicate via discrete spikes. Together, they promise a fully event-driven vision pipeline whose efficiency gains would benefit robots of any size---but this dissertation targets the hardest case, asking: \emph{how can a flying robot learn to navigate autonomously when it only has event-based vision?} Four research questions decompose this problem into a systematic progression from learning to deployment, adaptation, and sensor minimalism.

\cref{ch:neurips} establishes that spiking neural networks can learn complex visual tasks, presenting the first deep spiking networks to solve dense optical flow estimation from events in a self-supervised manner and demonstrating that spiking networks can move beyond simple classification to real-world vision problems relevant for robotics. \cref{ch:sr} takes the next step by deploying a fully neuromorphic vision-to-control pipeline on Intel's Loihi neuromorphic processor aboard a flying quadrotor, achieving the first autonomous drone flight with a fully neuromorphic system---running inference at 200~Hz consuming only 7--12~mW. \cref{ch:cvpr} brings learning on board by optimizing the self-supervised contrast maximization framework for computational efficiency, achieving a 100-fold runtime reduction that enables online depth learning during flight; just two minutes of on-device adaptation improved obstacle avoidance performance by approximately 30\% over just pre-training. Finally, \cref{ch:npjr} pushes toward ultimate sensor minimalism by demonstrating that vision can replace inertial measurement units entirely: a recurrent neural network learns to estimate attitude and rotation rates from events alone, enabling the first stable vision-only and IMU-free quadrotor flight.

Together, these contributions chart a path from theoretical promise to practical deployment for neuromorphic event-based robot vision, demonstrating that event-based vision is not merely a compelling concept but a viable foundation for autonomous flight. As the technology matures, it will enable a new generation of autonomous drones that perceive and navigate the world through efficient, brain-inspired vision.
